### 2009-02-17 presentation/presentation.out.tex

\documentclass[titlepage,usenames,a4,landscape,semhelv]{seminar}
\newcommand{\presenter}{Abram Hindle}
\newcommand{\conference}{}
\newcommand{\gettitle}{Temporal Developer Concept Witchery: Latent Dirichlet Allocation Applied to Revisions}

\newcommand{\gettitleproper}{\gettitle}
\newcommand{\names}{Abram Hindle, Micheal W. Godfrey, Richard C. Holt}
\author{
\names \\ 
{\small Software Architecture Group }\\
\small David R. Cheriton School of Computer Science\\
\small University of Waterloo\\
\small Canada\\
ahindle@cs.uwaterloo.ca
}

\include{header}

\newcommand{\imageslide}[2]{
\newslide
\includegraphics[width=0.9\textwidth]{#2}
}

\newcommand{\figcaption}[1]{
\begin{center}
#1\\
\end{center}
}

\newcommand{\ig}[1]{\includegraphics{#1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\pagestyle{fancy} %bars..
\begin{slide}

\begin{center}
{\bf \LARGE \gettitleproper }

{\names }

{\small Software Architecture Group }\\[-.5em]
{\small David R. Cheriton School of Computer Science}\\[-.5em]
{\small University of Waterloo}\\[-.5em]
{\small Canada}\\[-.5em]
{\small http://swag.uwaterloo.ca/}\\
\{ahindle,migod,holt\}@cs.uwaterloo.ca

\end{center}

\sslide{Introduction}
\begin{itemize}
\item Intro

\newslide

\begin{table*}
\centering
\begin{tabular}{|ccc|l|}

%\hline
%2000 &	Jul &	31 &		chmod \\
%2000 &	Sep &	29 &		fixes benchmark logging windows \\
%2000 &	Nov &	28 &		typo fix insert\_multi\_value \\
%2001 &	Jan &	27 &		fixes Innobase Cleanups auto-union \\
%2001 &	Mar &	28 &		2 topics bugfix, logging , TEMPORARY,	\\
\hline
2001 &	Jul &	26 &		update Allow TABLES LOCK [a] \\

2001 &	Aug &	25 &		tables row version [a] \\
\hline
%2001 &	Sep &	24 &		update checksum merge \\
%2001 &	Oct &	24 &		fixed fix \\
%2001 &	Dec &	23 &		HPUX SCO fix \\
\hline
2002 &	Feb &	21 &		net buffer length	max\_allowed\_packet [b] \\
2002 &	Mar &	23 &		small buf fix [b]	\\
%\hline
%2002 &	May &	22 &		[popular] fix SCO OSF1 table\_name \\
%2002 &	Nov &	18 &		HPUX11 compiler HP \\
\hline
\hline
2003 &	Feb &	16 &		Linux errno	[c] \\
2003 &	Mar &	18 &		alarm bookmark bug [c] \\
%\hline
%2003 &	Sep &	14 &		Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Continuous blocks found while tracking topics associated with the word portability in MySQL 3.23}
\label{tab:portability}
\end{table*}

\newslide

% \begin{figure}
%	 \centering
%	 \includegraphics[width=1.5\textwidth]{lda}
%	 \caption{Example of topics extracked from MySQL 3.23}
%	 \label{fig:mysql323}
% \end{figure}

\igbigslidecap{lda}{Example of topics extracted from MySQL 3.23}

\end{itemize}
\end{slide}

\end{document}