### 2010-01-27 icsm09/lda-lab.tex
moving older files into icsm09 directory

\documentclass{report}
\usepackage{graphicx,amsmath,comment,fullpage}
\newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
\newcommand{\igW}[1]{\includegraphics[width=.9\textwidth]{#1}}
\newcommand{\igWh}[1]{\includegraphics[width=.7\textwidth]{#1}}
\newcommand{\igWhalf}[1]{\includegraphics[width=.45\textwidth]{#1}}
\newcommand{\lda}{Latent Dirichlet Allocation}

\author{Abram Hindle\\
ahindle@cs.uwaterloo.ca\\
Software Architecture Group (SWAG)\\
Supervised by\\
Micheal Godfrey and Ric Holt
}
\title{Lab 0006: Latent Dirichlet Allocation Applied to Revisions}

\usepackage{rotating}   
\begin{document}

\maketitle
\begin{abstract}

  In this lab report we seek to explore how topics shift over time in
  the source control repository of a project. We analyze commit
  comments, and even source code per revision per each time window and
  extract the topics prevelant in that window. We expect that topic
  sets will change over time and the change in topic sets indicate a
  change in the focus of development. We suppose this could help us to
  segment the revisions on the time line to discover parts of an
  iteration. For topic extraction we use \lda (LDA).

\end{abstract}

\section{Introduction}

%?

In this lab report we seek to explore how topics shift over time in
the source control repository of a project. We analyze commit
comments, and even source code per revision per each time window and
extract the topics prevelant in that window. We expect that topic sets
will change over time and the change in topic sets indicate a change
in the focus of development. We suppose this could help us to segment
the revisions on the time line to discover parts of an iteration. For
topic extraction we use \lda (LDA).

\section{Methodology}

\subsection{Datasets and Tools}

Extractors:
\begin{itemize}
\item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
\item softChange - extract CVS facts to a PostgreSQL database.
\item bt2csv - Convert BitKeeper repositories to facts in CSV
  databases.
\end{itemize}

Analysis Tools:
\begin{itemize}
\item Hiraldo-Grok - An OCaml based spin off of Grok used for
  answering queries.
\item Gnuplot - graph plotting package
\end{itemize}

\section{First Impression}

% 2007-10-12

Looking at 30 day non-overlapping windows of revisions for MySQL 3.23
(20 topics, top 10 words per topic) we found there were common words
across topic clusters such as diffs, annotate and history. There were
notable transitional topics such as in the first window the word
bitkeeper appears (probably when they adopted bitkeeper) yet in the
following windows there were no mentions of bitkeeper. ``RENAME'' also
only appeared once in the first window.

Often words are shared across topics, see talbe \ref{tab:portability}
for tracking topics about portability.

It is apparent that we have to compare these topics some how by
itemset or similarity.

\begin{table}
\centering
\begin{tabular}{|ccc|l|}
\hline
2000 &  Jul &  31 &    chmod \\
2000 &  Sep &  29 &    fixes benchmark logging windows \\
2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
2001 &  Jan &  27 &    FIxes Innobase Cleanups auto-union \\
2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
\hline
2001 &  Jul &  26 &    Update Allow TABLES LOCK [a] \\

2001 &  Aug &  25 &    Tables row version [a] \\
\hline
2001 &  Sep &  24 &    Update checksum merge \\
2001 &  Oct &  24 &    fixed fix \\
2001 &  Dec &  23 &    HPUX SCO fix \\
\hline
2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
2002 &  Mar &  23 &    small buf fix [b]  \\
\hline
2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
2002 &  Nov &  18 &    HPUX11 compiler HP \\
\hline
2003 &  Feb &  16 &    Linux errno  [c] \\
2003 &  Mar &  18 &    alarm bookmark bug [c] \\
\hline
2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Tracking topics associated with the word portability, note some continuous blocks}
\label{tab:portability}
\end{table}

\end{document}