### 2010-02-27 src/validate/output/maxdb.mulan.out.transpose.tex
Mulan bug fixes and reporting!

\begin{tabular}{l|rrrrrrrrrrrrrrrrrr}

  & example-Accuracy & example-Fmeasure & example-HammingLoss & example-Precision & example-Recall & example-SubsetAccuracy & label-macro-AUC & label-macro-F1 & label-macro-Precision & label-macro-Recall & label-micro-AUC & label-micro-F1 & label-micro-Precision & label-micro-Recall & rank-AvgPrecision & rank-Coverage & rank-One-error & rank-Ranking Loss \\

\hline \\

BR & 0.49 & 0.55 & 0.19 & 0.50 & 0.61 & 0.40 & NaN & 0.32 & 0.29 & 0.38 & 0.76 & 0.39 & 0.34 & 0.47 & 0.54 & 1.40 & 0.81 & 0.41 \\

CLR & 0.47 & 0.50 & 0.14 & 0.49 & 0.50 & 0.43 & NaN & 0.19 & 0.31 & 0.15 & 0.66 & 0.27 & 0.42 & 0.21 & 0.57 & 1.23 & 0.79 & 0.35 \\

HOMER & 0.51 & 0.53 & 0.14 & 0.53 & 0.53 & 0.45 & NaN & 0.23 & 0.32 & 0.20 & 0.62 & 0.34 & 0.49 & 0.26 & 0.54 & 1.43 & 0.81 & 0.41 \\

IncludeLabels & 0.44 & 0.45 & 0.14 & 0.45 & 0.45 & 0.43 & NaN & 0.11 & 0.21 & 0.08 & 0.55 & 0.13 & 0.33 & 0.09 & 0.40 & 1.84 & 0.90 & 0.58 \\

LP & 0.42 & 0.44 & 0.17 & 0.45 & 0.43 & 0.38 & NaN & 0.19 & 0.23 & 0.17 & 0.62 & 0.25 & 0.30 & 0.22 & 0.40 & 1.93 & 0.89 & 0.60 \\

MLStacking & 0.47 & 0.48 & 0.13 & 0.49 & 0.47 & 0.43 & NaN & 0.12 & 0.18 & 0.10 & 0.68 & 0.21 & 0.49 & 0.14 & 0.60 & 1.16 & 0.76 & 0.32 \\

MLkNN & 0.43 & 0.43 & 0.13 & 0.43 & 0.43 & 0.43 & NaN & 0.00 & 0.00 & 0.00 & 0.65 & 0.00 & 0.00 & 0.00 & 0.58 & 1.20 & 0.79 & 0.33 \\

RAkEL & 0.48 & 0.50 & 0.13 & 0.51 & 0.50 & 0.43 & NaN & 0.26 & 0.40 & 0.23 & 0.68 & 0.34 & 0.49 & 0.27 & 0.51 & 1.54 & 0.79 & 0.46 \\

\end{tabular}