### 2010-03-02 src/validate/output/maxdb.mulan.out.tex

\begin{tabular}{l|rrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\

\hline \\

example-Accuracy & 0.49 & 0.47 & 0.51 & 0.44 & 0.42 & 0.47 & 0.43 & 0.48 \\

example-Fmeasure & 0.55 & 0.50 & 0.53 & 0.45 & 0.44 & 0.48 & 0.43 & 0.50 \\

example-HammingLoss & 0.19 & 0.14 & 0.14 & 0.14 & 0.17 & 0.13 & 0.13 & 0.13 \\

example-Precision & 0.50 & 0.49 & 0.53 & 0.45 & 0.45 & 0.49 & 0.43 & 0.51 \\

example-Recall & 0.61 & 0.50 & 0.53 & 0.45 & 0.43 & 0.47 & 0.43 & 0.50 \\

example-SubsetAccuracy & 0.40 & 0.43 & 0.45 & 0.43 & 0.38 & 0.43 & 0.43 & 0.43 \\

label-macro-AUC & NaN & NaN & NaN & NaN & NaN & NaN & NaN & NaN \\

label-macro-F1 & 0.32 & 0.19 & 0.23 & 0.11 & 0.19 & 0.12 & 0.00 & 0.26 \\

label-macro-Precision & 0.29 & 0.31 & 0.32 & 0.21 & 0.23 & 0.18 & 0.00 & 0.40 \\

label-macro-Recall & 0.38 & 0.15 & 0.20 & 0.08 & 0.17 & 0.10 & 0.00 & 0.23 \\

label-micro-AUC & 0.76 & 0.66 & 0.62 & 0.55 & 0.62 & 0.68 & 0.65 & 0.68 \\

label-micro-F1 & 0.39 & 0.27 & 0.34 & 0.13 & 0.25 & 0.21 & 0.00 & 0.34 \\

label-micro-Precision & 0.34 & 0.42 & 0.49 & 0.33 & 0.30 & 0.49 & 0.00 & 0.49 \\

label-micro-Recall & 0.47 & 0.21 & 0.26 & 0.09 & 0.22 & 0.14 & 0.00 & 0.27 \\

rank-AvgPrecision & 0.54 & 0.57 & 0.54 & 0.40 & 0.40 & 0.60 & 0.58 & 0.51 \\

rank-Coverage & 1.40 & 1.23 & 1.43 & 1.84 & 1.93 & 1.16 & 1.20 & 1.54 \\

rank-One-error & 0.81 & 0.79 & 0.81 & 0.90 & 0.89 & 0.76 & 0.79 & 0.79 \\

rank-Ranking Loss & 0.41 & 0.35 & 0.41 & 0.58 & 0.60 & 0.32 & 0.33 & 0.46 \\

\end{tabular}