### 2011-08-25 src/validate/output/maxdb.mulan.out.tex

\begin{tabular}{l|rrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\

\hline \\

example-Accuracy & 0.50 & 0.54 & 0.58 & 0.56 & 0.59 & 0.53 & 0.43 & 0.54 \\

example-Fmeasure & 0.56 & 0.58 & 0.60 & 0.59 & 0.62 & 0.55 & 0.43 & 0.56 \\

example-HammingLoss & 0.19 & 0.14 & 0.13 & 0.14 & 0.17 & 0.14 & 0.13 & 0.14 \\

example-Precision & 0.50 & 0.57 & 0.61 & 0.59 & 0.62 & 0.57 & 0.43 & 0.57 \\

example-Recall & 0.62 & 0.59 & 0.60 & 0.59 & 0.61 & 0.53 & 0.43 & 0.55 \\

example-SubsetAccuracy & 0.41 & 0.48 & 0.52 & 0.50 & 0.53 & 0.48 & 0.43 & 0.49 \\

label-macro-AUC & NaN & NaN & NaN & NaN & NaN & NaN & NaN & NaN \\

label-macro-F1 & 0.31 & 0.26 & 0.25 & 0.22 & 0.25 & 0.11 & 0.00 & 0.25 \\

label-macro-Precision & 0.29 & 0.32 & 0.31 & 0.29 & 0.29 & 0.16 & 0.02 & 0.34 \\

label-macro-Recall & 0.39 & 0.24 & 0.23 & 0.21 & 0.26 & 0.12 & 0.00 & 0.22 \\

label-micro-AUC & 0.77 & 0.67 & 0.61 & 0.77 & 0.74 & 0.72 & 0.67 & 0.74 \\

label-micro-F1 & 0.39 & 0.36 & 0.38 & 0.34 & 0.33 & 0.26 & 0.01 & 0.33 \\

label-micro-Precision & 0.34 & 0.44 & 0.48 & 0.44 & 0.35 & 0.49 & 0.10 & 0.47 \\

label-micro-Recall & 0.47 & 0.31 & 0.32 & 0.29 & 0.32 & 0.19 & 0.00 & 0.27 \\

rank-AvgPrecision & 0.53 & 0.59 & 0.54 & 0.56 & 0.47 & 0.60 & 0.56 & 0.56 \\

rank-Coverage & 1.41 & 1.19 & 1.41 & 1.33 & 1.82 & 1.16 & 1.24 & 1.40 \\

rank-One-error & 0.83 & 0.78 & 0.81 & 0.79 & 0.83 & 0.77 & 0.82 & 0.77 \\

rank-Ranking Loss & 0.41 & 0.34 & 0.41 & 0.38 & 0.55 & 0.31 & 0.35 & 0.39 \\

\end{tabular}