### 2011-08-25 src/validate/output/maxdb.mulan.out.transpose.tex

\begin{tabular}{l|rrrrrrrrrrrrrrrrrr}

  & example-Accuracy & example-Fmeasure & example-HammingLoss & example-Precision & example-Recall & example-SubsetAccuracy & label-macro-AUC & label-macro-F1 & label-macro-Precision & label-macro-Recall & label-micro-AUC & label-micro-F1 & label-micro-Precision & label-micro-Recall & rank-AvgPrecision & rank-Coverage & rank-One-error & rank-Ranking Loss \\

\hline \\

BR & 0.50 & 0.56 & 0.19 & 0.50 & 0.62 & 0.41 & NaN & 0.31 & 0.29 & 0.39 & 0.77 & 0.39 & 0.34 & 0.47 & 0.53 & 1.41 & 0.83 & 0.41 \\

CLR & 0.54 & 0.58 & 0.14 & 0.57 & 0.59 & 0.48 & NaN & 0.26 & 0.32 & 0.24 & 0.67 & 0.36 & 0.44 & 0.31 & 0.59 & 1.19 & 0.78 & 0.34 \\

HOMER & 0.58 & 0.60 & 0.13 & 0.61 & 0.60 & 0.52 & NaN & 0.25 & 0.31 & 0.23 & 0.61 & 0.38 & 0.48 & 0.32 & 0.54 & 1.41 & 0.81 & 0.41 \\

IncludeLabels & 0.56 & 0.59 & 0.14 & 0.59 & 0.59 & 0.50 & NaN & 0.22 & 0.29 & 0.21 & 0.77 & 0.34 & 0.44 & 0.29 & 0.56 & 1.33 & 0.79 & 0.38 \\

LP & 0.59 & 0.62 & 0.17 & 0.62 & 0.61 & 0.53 & NaN & 0.25 & 0.29 & 0.26 & 0.74 & 0.33 & 0.35 & 0.32 & 0.47 & 1.82 & 0.83 & 0.55 \\

MLStacking & 0.53 & 0.55 & 0.14 & 0.57 & 0.53 & 0.48 & NaN & 0.11 & 0.16 & 0.12 & 0.72 & 0.26 & 0.49 & 0.19 & 0.60 & 1.16 & 0.77 & 0.31 \\

MLkNN & 0.43 & 0.43 & 0.13 & 0.43 & 0.43 & 0.43 & NaN & 0.00 & 0.02 & 0.00 & 0.67 & 0.01 & 0.10 & 0.00 & 0.56 & 1.24 & 0.82 & 0.35 \\

RAkEL & 0.54 & 0.56 & 0.14 & 0.57 & 0.55 & 0.49 & NaN & 0.25 & 0.34 & 0.22 & 0.74 & 0.33 & 0.47 & 0.27 & 0.56 & 1.40 & 0.77 & 0.39 \\

\end{tabular}