### 2011-08-25 src/validate/output/mysql.mulan.out.tex

\begin{tabular}{l|rrrr}

  & BR & CLR & IncludeLabels & MLkNN \\

\hline \\

example-Accuracy & 0.46 & 0.41 & 0.41 & 0.38 \\

example-Fmeasure & 0.59 & 0.50 & 0.49 & 0.46 \\

example-HammingLoss & 0.25 & 0.21 & 0.21 & 0.21 \\

example-Precision & 0.51 & 0.52 & 0.53 & 0.50 \\

example-Recall & 0.68 & 0.48 & 0.46 & 0.42 \\

example-SubsetAccuracy & 0.18 & 0.22 & 0.23 & 0.23 \\

label-macro-AUC & 0.74 & 0.65 & 0.61 & 0.61 \\

label-macro-F1 & 0.47 & 0.34 & 0.31 & 0.25 \\

label-macro-Precision & 0.42 & 0.41 & 0.41 & 0.33 \\

label-macro-Recall & 0.57 & 0.30 & 0.27 & 0.22 \\

label-micro-AUC & 0.81 & 0.80 & 0.77 & 0.78 \\

label-micro-F1 & 0.59 & 0.54 & 0.52 & 0.49 \\

label-micro-Precision & 0.52 & 0.65 & 0.67 & 0.69 \\

label-micro-Recall & 0.69 & 0.47 & 0.43 & 0.39 \\

rank-AvgPrecision & 0.76 & 0.76 & 0.73 & 0.74 \\

rank-Coverage & 1.67 & 1.64 & 1.87 & 1.73 \\

rank-One-error & 0.39 & 0.38 & 0.41 & 0.40 \\

rank-Ranking Loss & 0.19 & 0.19 & 0.23 & 0.21 \\

\end{tabular}