### 2011-08-25 src/validate/output/mysql.mulan.out.transpose.tex

\begin{tabular}{l|rrrrrrrrrrrrrrrrrr}

  & example-Accuracy & example-Fmeasure & example-HammingLoss & example-Precision & example-Recall & example-SubsetAccuracy & label-macro-AUC & label-macro-F1 & label-macro-Precision & label-macro-Recall & label-micro-AUC & label-micro-F1 & label-micro-Precision & label-micro-Recall & rank-AvgPrecision & rank-Coverage & rank-One-error & rank-Ranking Loss \\

\hline \\

BR & 0.46 & 0.59 & 0.25 & 0.51 & 0.68 & 0.18 & 0.74 & 0.47 & 0.42 & 0.57 & 0.81 & 0.59 & 0.52 & 0.69 & 0.76 & 1.67 & 0.39 & 0.19 \\

CLR & 0.41 & 0.50 & 0.21 & 0.52 & 0.48 & 0.22 & 0.65 & 0.34 & 0.41 & 0.30 & 0.80 & 0.54 & 0.65 & 0.47 & 0.76 & 1.64 & 0.38 & 0.19 \\

IncludeLabels & 0.41 & 0.49 & 0.21 & 0.53 & 0.46 & 0.23 & 0.61 & 0.31 & 0.41 & 0.27 & 0.77 & 0.52 & 0.67 & 0.43 & 0.73 & 1.87 & 0.41 & 0.23 \\

MLkNN & 0.38 & 0.46 & 0.21 & 0.50 & 0.42 & 0.23 & 0.61 & 0.25 & 0.33 & 0.22 & 0.78 & 0.49 & 0.69 & 0.39 & 0.74 & 1.73 & 0.40 & 0.21 \\

\end{tabular}