### 2011-09-01 src/validate/output/mysql.mulan.out.tex
Updating to a newer version of mulan

\begin{tabular}{l|}

  \\

\hline \\

\end{tabular}