### 2011-09-02 src/validate/output/maxdb.mulan.out.tex
Mulan parser and new run of data

\begin{tabular}{l|rrrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MC-Copy & MC-Ignore & MLkNN & RAkEL \\

\hline \\

Average Precision & 0.53 & 0.59 & 0.52 & 0.56 & 0.47 & 0.52 & 0.55 & 0.58 & 0.54 \\

Coverage & 1.41 & 1.19 & 1.44 & 1.33 & 1.82 & 1.52 & 1.54 & 1.21 & 1.49 \\

ErrorSetSize & 1.42 & 1.13 & 1.38 & 1.29 & 1.87 & 1.49 & 1.54 & 1.14 & 1.43 \\

Example-Based Accuracy & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based F Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Hamming Loss & 0.19 & 0.14 & 0.13 & 0.14 & 0.17 &  &  & 0.13 & 0.14 \\

IsError & 0.44 & 0.41 & 0.45 & 0.42 & 0.46 & 0.44 & 0.41 & 0.42 & 0.42 \\

Macro-averaged AUC & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged F-Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Mean Average Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Micro-averaged AUC & 0.77 & 0.67 & 0.61 & 0.77 & 0.74 &  &  & 0.69 & 0.72 \\

Micro-averaged F-Measure & 0.39 & 0.36 & 0.37 & 0.34 & 0.33 &  &  & NaN & 0.34 \\

Micro-averaged Precision & 0.34 & 0.44 & 0.52 & 0.44 & 0.35 &  &  & NaN & 0.44 \\

Micro-averaged Recall & 0.47 & 0.31 & 0.30 & 0.29 & 0.32 &  &  & 0.00 & 0.28 \\

OneError & 0.83 & 0.78 & 0.83 & 0.79 & 0.83 & 0.81 & 0.78 & 0.79 & 0.79 \\

Ranking Loss & 0.23 & 0.19 & 0.24 & 0.22 & 0.31 &  &  & 0.19 & 0.24 \\

Subset Accuracy & 0.41 & 0.48 & 0.50 & 0.50 & 0.53 &  &  & 0.43 & 0.51 \\

\end{tabular}