### 2011-09-02 src/validate/output/maxdb.mulan.out.transpose.tex

\begin{tabular}{l|rrrrrrrrrrrrrrrrrrrrr}

  & Average Precision & Coverage & ErrorSetSize & Example-Based Accuracy & Example-Based F Measure & Example-Based Precision & Example-Based Recall & Hamming Loss & IsError & Macro-averaged AUC & Macro-averaged F-Measure & Macro-averaged Precision & Macro-averaged Recall & Mean Average Precision & Micro-averaged AUC & Micro-averaged F-Measure & Micro-averaged Precision & Micro-averaged Recall & OneError & Ranking Loss & Subset Accuracy \\

\hline \\

BR & 0.53 & 1.41 & 1.42 & NaN & NaN & NaN & NaN & 0.19 & 0.44 & NaN & NaN & NaN & NaN & NaN & 0.77 & 0.39 & 0.34 & 0.47 & 0.83 & 0.23 & 0.41 \\

CLR & 0.59 & 1.19 & 1.13 & NaN & NaN & NaN & NaN & 0.14 & 0.41 & NaN & NaN & NaN & NaN & NaN & 0.67 & 0.36 & 0.44 & 0.31 & 0.78 & 0.19 & 0.48 \\

HOMER & 0.52 & 1.44 & 1.38 & NaN & NaN & NaN & NaN & 0.13 & 0.45 & NaN & NaN & NaN & NaN & NaN & 0.61 & 0.37 & 0.52 & 0.30 & 0.83 & 0.24 & 0.50 \\

IncludeLabels & 0.56 & 1.33 & 1.29 & NaN & NaN & NaN & NaN & 0.14 & 0.42 & NaN & NaN & NaN & NaN & NaN & 0.77 & 0.34 & 0.44 & 0.29 & 0.79 & 0.22 & 0.50 \\

LP & 0.47 & 1.82 & 1.87 & NaN & NaN & NaN & NaN & 0.17 & 0.46 & NaN & NaN & NaN & NaN & NaN & 0.74 & 0.33 & 0.35 & 0.32 & 0.83 & 0.31 & 0.53 \\

MC-Copy & 0.52 & 1.52 & 1.49 &  &  &  &  &  & 0.44 &  &  &  &  &  &  &  &  &  & 0.81 &  &  \\

MC-Ignore & 0.55 & 1.54 & 1.54 &  &  &  &  &  & 0.41 &  &  &  &  &  &  &  &  &  & 0.78 &  &  \\

MLkNN & 0.58 & 1.21 & 1.14 & NaN & NaN & NaN & NaN & 0.13 & 0.42 & NaN & NaN & NaN & NaN & NaN & 0.69 & NaN & NaN & 0.00 & 0.79 & 0.19 & 0.43 \\

RAkEL & 0.54 & 1.49 & 1.43 & NaN & NaN & NaN & NaN & 0.14 & 0.42 & NaN & NaN & NaN & NaN & NaN & 0.72 & 0.34 & 0.44 & 0.28 & 0.79 & 0.24 & 0.51 \\

\end{tabular}