### 2011-09-02 src/validate/output/mysql.mulan.out.tex

\begin{tabular}{l|rrrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MC-Copy & MC-Ignore & MLkNN & RAkEL \\

\hline \\

Average Precision & 0.76 & 0.76 & 0.69 & 0.73 & 0.53 & 0.72 & 0.68 & 0.75 & 0.74 \\

Coverage & 1.67 & 1.64 & 2.24 & 1.87 & 3.14 & 1.88 & 2.35 & 1.69 & 2.01 \\

ErrorSetSize & 1.23 & 1.18 & 1.90 & 1.48 & 3.49 & 1.49 & 2.10 & 1.25 & 1.59 \\

Example-Based Accuracy & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based F Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Hamming Loss & 0.25 & 0.21 & 0.20 & 0.21 & 0.25 &  &  & 0.22 & 0.20 \\

IsError & 0.55 & 0.52 & 0.62 & 0.57 & 0.74 & 0.57 & 0.64 & 0.54 & 0.55 \\

Macro-averaged AUC & 0.74 & 0.65 & 0.64 & 0.61 & 0.59 &  &  & 0.62 & 0.68 \\

Macro-averaged F-Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Precision & 0.42 & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Recall & 0.57 & 0.30 & 0.45 & 0.27 & 0.33 &  &  & 0.22 & 0.34 \\

Mean Average Precision & 0.53 & 0.44 & 0.45 & 0.42 & 0.39 &  &  & 0.41 & 0.50 \\

Micro-averaged AUC & 0.81 & 0.80 & 0.74 & 0.77 & 0.66 &  &  & 0.79 & 0.80 \\

Micro-averaged F-Measure & 0.59 & 0.54 & 0.61 & 0.52 & 0.50 &  &  & 0.48 & 0.56 \\

Micro-averaged Precision & 0.52 & 0.65 & 0.63 & 0.67 & 0.53 &  &  & 0.67 & 0.66 \\

Micro-averaged Recall & 0.69 & 0.47 & 0.59 & 0.43 & 0.46 &  &  & 0.38 & 0.49 \\

OneError & 0.39 & 0.38 & 0.45 & 0.41 & 0.62 & 0.42 & 0.44 & 0.39 & 0.37 \\

Ranking Loss & 0.18 & 0.18 & 0.28 & 0.22 & 0.50 &  &  & 0.19 & 0.24 \\

Subset Accuracy & 0.18 & 0.22 & 0.30 & 0.23 & 0.25 &  &  & 0.22 & 0.29 \\

\end{tabular}