### 2011-09-02 src/validate/output/mysql.mulan.out.transpose.tex

\begin{tabular}{l|rrrrrrrrrrrrrrrrrrrrr}

  & Average Precision & Coverage & ErrorSetSize & Example-Based Accuracy & Example-Based F Measure & Example-Based Precision & Example-Based Recall & Hamming Loss & IsError & Macro-averaged AUC & Macro-averaged F-Measure & Macro-averaged Precision & Macro-averaged Recall & Mean Average Precision & Micro-averaged AUC & Micro-averaged F-Measure & Micro-averaged Precision & Micro-averaged Recall & OneError & Ranking Loss & Subset Accuracy \\

\hline \\

BR & 0.76 & 1.67 & 1.23 & NaN & NaN & NaN & NaN & 0.25 & 0.55 & 0.74 & NaN & 0.42 & 0.57 & 0.53 & 0.81 & 0.59 & 0.52 & 0.69 & 0.39 & 0.18 & 0.18 \\

CLR & 0.76 & 1.64 & 1.18 & NaN & NaN & NaN & NaN & 0.21 & 0.52 & 0.65 & NaN & NaN & 0.30 & 0.44 & 0.80 & 0.54 & 0.65 & 0.47 & 0.38 & 0.18 & 0.22 \\

HOMER & 0.69 & 2.24 & 1.90 & NaN & NaN & NaN & NaN & 0.20 & 0.62 & 0.64 & NaN & NaN & 0.45 & 0.45 & 0.74 & 0.61 & 0.63 & 0.59 & 0.45 & 0.28 & 0.30 \\

IncludeLabels & 0.73 & 1.87 & 1.48 & NaN & NaN & NaN & NaN & 0.21 & 0.57 & 0.61 & NaN & NaN & 0.27 & 0.42 & 0.77 & 0.52 & 0.67 & 0.43 & 0.41 & 0.22 & 0.23 \\

LP & 0.53 & 3.14 & 3.49 & NaN & NaN & NaN & NaN & 0.25 & 0.74 & 0.59 & NaN & NaN & 0.33 & 0.39 & 0.66 & 0.50 & 0.53 & 0.46 & 0.62 & 0.50 & 0.25 \\

MC-Copy & 0.72 & 1.88 & 1.49 &  &  &  &  &  & 0.57 &  &  &  &  &  &  &  &  &  & 0.42 &  &  \\

MC-Ignore & 0.68 & 2.35 & 2.10 &  &  &  &  &  & 0.64 &  &  &  &  &  &  &  &  &  & 0.44 &  &  \\

MLkNN & 0.75 & 1.69 & 1.25 & NaN & NaN & NaN & NaN & 0.22 & 0.54 & 0.62 & NaN & NaN & 0.22 & 0.41 & 0.79 & 0.48 & 0.67 & 0.38 & 0.39 & 0.19 & 0.22 \\

RAkEL & 0.74 & 2.01 & 1.59 & NaN & NaN & NaN & NaN & 0.20 & 0.55 & 0.68 & NaN & NaN & 0.34 & 0.50 & 0.80 & 0.56 & 0.66 & 0.49 & 0.37 & 0.24 & 0.29 \\

\end{tabular}