### 2011-09-02 src/validate/output/pgsqla.mulan.out.tex

\begin{tabular}{l|rrrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MC-Copy & MC-Ignore & MLkNN & RAkEL \\

\hline \\

Average Precision & 0.78 & 0.78 & 0.74 & 0.69 & 0.59 & 0.83 & 0.68 & 0.85 & 0.80 \\

Coverage & 2.06 & 2.01 & 2.35 & 2.66 & 3.30 & 1.83 & 3.00 & 1.74 & 2.15 \\

ErrorSetSize & 1.48 & 1.40 & 1.81 & 2.29 & 3.49 & 1.16 & 2.72 & 1.02 & 1.52 \\

Example-Based Accuracy & 0.51 & NaN & 0.58 & NaN & NaN &  &  & 0.56 & NaN \\

Example-Based F Measure & NaN & NaN & 0.75 & NaN & NaN &  &  & 0.70 & 0.70 \\

Example-Based Precision & NaN & NaN & 0.71 & NaN & NaN &  &  & NaN & NaN \\

Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Hamming Loss & 0.25 & 0.25 & 0.20 & 0.26 & 0.25 &  &  & 0.20 & 0.21 \\

IsError & 0.58 & 0.60 & 0.62 & 0.69 & 0.76 & 0.47 & 0.67 & 0.45 & 0.53 \\

Macro-averaged AUC & 0.63 & 0.56 & 0.54 & 0.51 & 0.54 &  &  & 0.54 & 0.53 \\

Macro-averaged F-Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Recall & 0.46 & 0.37 & 0.38 & 0.32 & 0.37 &  &  & 0.29 & 0.33 \\

Mean Average Precision & 0.45 & 0.40 & 0.39 & 0.39 & 0.39 &  &  & 0.38 & 0.39 \\

Micro-averaged AUC & 0.81 & 0.81 & 0.79 & 0.70 & 0.73 &  &  & 0.84 & 0.80 \\

Micro-averaged F-Measure & 0.64 & 0.62 & 0.69 & 0.58 & 0.60 &  &  & 0.67 & 0.66 \\

Micro-averaged Precision & 0.60 & 0.60 & 0.68 & 0.60 & 0.60 &  &  & 0.72 & 0.67 \\

Micro-averaged Recall & 0.69 & 0.64 & 0.70 & 0.57 & 0.61 &  &  & 0.63 & 0.64 \\

OneError & 0.31 & 0.28 & 0.38 & 0.42 & 0.57 & 0.21 & 0.36 & 0.19 & 0.25 \\

Ranking Loss & 0.20 & 0.19 & 0.24 & 0.31 & 0.46 &  &  & 0.14 & 0.20 \\

Subset Accuracy & 0.19 & 0.17 & 0.27 & 0.18 & 0.20 &  &  & 0.28 & 0.26 \\

\end{tabular}