### 2011-09-02 src/validate/output/pgsqla.mulan.out.transpose.tex

\begin{tabular}{l|rrrrrrrrrrrrrrrrrrrrr}

  & Average Precision & Coverage & ErrorSetSize & Example-Based Accuracy & Example-Based F Measure & Example-Based Precision & Example-Based Recall & Hamming Loss & IsError & Macro-averaged AUC & Macro-averaged F-Measure & Macro-averaged Precision & Macro-averaged Recall & Mean Average Precision & Micro-averaged AUC & Micro-averaged F-Measure & Micro-averaged Precision & Micro-averaged Recall & OneError & Ranking Loss & Subset Accuracy \\

\hline \\

BR & 0.78 & 2.06 & 1.48 & 0.51 & NaN & NaN & NaN & 0.25 & 0.58 & 0.63 & NaN & NaN & 0.46 & 0.45 & 0.81 & 0.64 & 0.60 & 0.69 & 0.31 & 0.20 & 0.19 \\

CLR & 0.78 & 2.01 & 1.40 & NaN & NaN & NaN & NaN & 0.25 & 0.60 & 0.56 & NaN & NaN & 0.37 & 0.40 & 0.81 & 0.62 & 0.60 & 0.64 & 0.28 & 0.19 & 0.17 \\

HOMER & 0.74 & 2.35 & 1.81 & 0.58 & 0.75 & 0.71 & NaN & 0.20 & 0.62 & 0.54 & NaN & NaN & 0.38 & 0.39 & 0.79 & 0.69 & 0.68 & 0.70 & 0.38 & 0.24 & 0.27 \\

IncludeLabels & 0.69 & 2.66 & 2.29 & NaN & NaN & NaN & NaN & 0.26 & 0.69 & 0.51 & NaN & NaN & 0.32 & 0.39 & 0.70 & 0.58 & 0.60 & 0.57 & 0.42 & 0.31 & 0.18 \\

LP & 0.59 & 3.30 & 3.49 & NaN & NaN & NaN & NaN & 0.25 & 0.76 & 0.54 & NaN & NaN & 0.37 & 0.39 & 0.73 & 0.60 & 0.60 & 0.61 & 0.57 & 0.46 & 0.20 \\

MC-Copy & 0.83 & 1.83 & 1.16 &  &  &  &  &  & 0.47 &  &  &  &  &  &  &  &  &  & 0.21 &  &  \\

MC-Ignore & 0.68 & 3.00 & 2.72 &  &  &  &  &  & 0.67 &  &  &  &  &  &  &  &  &  & 0.36 &  &  \\

MLkNN & 0.85 & 1.74 & 1.02 & 0.56 & 0.70 & NaN & NaN & 0.20 & 0.45 & 0.54 & NaN & NaN & 0.29 & 0.38 & 0.84 & 0.67 & 0.72 & 0.63 & 0.19 & 0.14 & 0.28 \\

RAkEL & 0.80 & 2.15 & 1.52 & NaN & 0.70 & NaN & NaN & 0.21 & 0.53 & 0.53 & NaN & NaN & 0.33 & 0.39 & 0.80 & 0.66 & 0.67 & 0.64 & 0.25 & 0.20 & 0.26 \\

\end{tabular}