### 2011-09-02 src/validate/output/pgsqln.mulan.out.tex

\begin{tabular}{l|rrrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MC-Copy & MC-Ignore & MLkNN & RAkEL \\

\hline \\

Average Precision & 0.60 & 0.56 & 0.56 & 0.53 & 0.47 & 0.53 & 0.52 & 0.57 & 0.55 \\

Coverage & 2.49 & 2.67 & 2.78 & 2.94 & 3.26 & 2.99 & 3.02 & 2.71 & 2.86 \\

ErrorSetSize & 2.41 & 2.67 & 2.74 & 3.01 & 3.57 & 3.07 & 3.18 & 2.68 & 2.86 \\

Example-Based Accuracy & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based F Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Hamming Loss & 0.34 & 0.33 & 0.32 & 0.34 & 0.36 &  &  & 0.27 & 0.30 \\

IsError & 0.74 & 0.77 & 0.80 & 0.82 & 0.87 & 0.80 & 0.81 & 0.77 & 0.81 \\

Macro-averaged AUC & 0.64 & 0.58 & 0.58 & 0.55 & 0.53 &  &  & 0.53 & 0.56 \\

Macro-averaged F-Measure & 0.41 & NaN & 0.37 & NaN & 0.28 &  &  & NaN & NaN \\

Macro-averaged Precision & 0.38 & 0.37 & 0.39 & 0.33 & 0.30 &  &  & NaN & 0.30 \\

Macro-averaged Recall & 0.48 & 0.34 & 0.39 & 0.31 & 0.29 &  &  & 0.00 & 0.22 \\

Mean Average Precision & 0.43 & 0.38 & 0.39 & 0.35 & 0.33 &  &  & 0.34 & 0.36 \\

Micro-averaged AUC & 0.65 & 0.60 & 0.60 & 0.56 & 0.55 &  &  & 0.59 & 0.58 \\

Micro-averaged F-Measure & 0.43 & 0.36 & 0.41 & 0.34 & 0.31 &  &  & NaN & 0.30 \\

Micro-averaged Precision & 0.39 & 0.38 & 0.40 & 0.35 & 0.31 &  &  & NaN & 0.39 \\

Micro-averaged Recall & 0.49 & 0.35 & 0.42 & 0.33 & 0.30 &  &  & 0.01 & 0.24 \\

OneError & 0.59 & 0.63 & 0.64 & 0.68 & 0.75 & 0.66 & 0.66 & 0.62 & 0.64 \\

Ranking Loss & 0.35 & 0.39 & 0.40 & 0.44 & 0.52 &  &  & 0.40 & 0.42 \\

Subset Accuracy & 0.08 & 0.10 & 0.10 & 0.09 & 0.11 &  &  & 0.04 & 0.10 \\

\end{tabular}