### 2011-09-02 src/validate/output/pgsqln.mulan.out.transpose.tex

\begin{tabular}{l|rrrrrrrrrrrrrrrrrrrrr}

  & Average Precision & Coverage & ErrorSetSize & Example-Based Accuracy & Example-Based F Measure & Example-Based Precision & Example-Based Recall & Hamming Loss & IsError & Macro-averaged AUC & Macro-averaged F-Measure & Macro-averaged Precision & Macro-averaged Recall & Mean Average Precision & Micro-averaged AUC & Micro-averaged F-Measure & Micro-averaged Precision & Micro-averaged Recall & OneError & Ranking Loss & Subset Accuracy \\

\hline \\

BR & 0.60 & 2.49 & 2.41 & NaN & NaN & NaN & NaN & 0.34 & 0.74 & 0.64 & 0.41 & 0.38 & 0.48 & 0.43 & 0.65 & 0.43 & 0.39 & 0.49 & 0.59 & 0.35 & 0.08 \\

CLR & 0.56 & 2.67 & 2.67 & NaN & NaN & NaN & NaN & 0.33 & 0.77 & 0.58 & NaN & 0.37 & 0.34 & 0.38 & 0.60 & 0.36 & 0.38 & 0.35 & 0.63 & 0.39 & 0.10 \\

HOMER & 0.56 & 2.78 & 2.74 & NaN & NaN & NaN & NaN & 0.32 & 0.80 & 0.58 & 0.37 & 0.39 & 0.39 & 0.39 & 0.60 & 0.41 & 0.40 & 0.42 & 0.64 & 0.40 & 0.10 \\

IncludeLabels & 0.53 & 2.94 & 3.01 & NaN & NaN & NaN & NaN & 0.34 & 0.82 & 0.55 & NaN & 0.33 & 0.31 & 0.35 & 0.56 & 0.34 & 0.35 & 0.33 & 0.68 & 0.44 & 0.09 \\

LP & 0.47 & 3.26 & 3.57 & NaN & NaN & NaN & NaN & 0.36 & 0.87 & 0.53 & 0.28 & 0.30 & 0.29 & 0.33 & 0.55 & 0.31 & 0.31 & 0.30 & 0.75 & 0.52 & 0.11 \\

MC-Copy & 0.53 & 2.99 & 3.07 &  &  &  &  &  & 0.80 &  &  &  &  &  &  &  &  &  & 0.66 &  &  \\

MC-Ignore & 0.52 & 3.02 & 3.18 &  &  &  &  &  & 0.81 &  &  &  &  &  &  &  &  &  & 0.66 &  &  \\

MLkNN & 0.57 & 2.71 & 2.68 & NaN & NaN & NaN & NaN & 0.27 & 0.77 & 0.53 & NaN & NaN & 0.00 & 0.34 & 0.59 & NaN & NaN & 0.01 & 0.62 & 0.40 & 0.04 \\

RAkEL & 0.55 & 2.86 & 2.86 & NaN & NaN & NaN & NaN & 0.30 & 0.81 & 0.56 & NaN & 0.30 & 0.22 & 0.36 & 0.58 & 0.30 & 0.39 & 0.24 & 0.64 & 0.42 & 0.10 \\

\end{tabular}